Inferring Continuous and Discrete Population Genetic Structure Across Space

### 2017-07-30 writeup/conStruct.tex
added to x-validation

\paragraph{Choice of cluster number and cross-validation}
There are a number of reasons, both statistical and biological, 
why choosing a single, ``best" number of clusters to describe the data may not be desirable.
%, including ambiguity in what exactly it means for any given value of $K$ to be ``best."
We discuss these reasons further in the Discussion below, 
but we also recognize that in many circumstances it is either useful or convenient 
to present only a single set of results from a clustering analysis with a particular number of clusters.
ideally a number large enough to describe meaningful partitions in the data, 
but not so large as to be difficult to parse visually.
Toward that end, we have implemented a method for 
statistically comparing \texttt{conStruct} runs across different values of $K$.

Several approaches have been proposed to determine the number of clusters in the data, 
including: 
the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
comparisons of the likelihood of the data across different values of $K$, 
with various criteria on how to choose a single value (e.g., Evanno 2005);
or comparisons of the marginal likelihood, 
generated either via various approximations \citep[e.g.][]{STRUCTURE}
or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
but see \cite{verity_nichols2016} for an excellent discussion of and comparison between several.

We choose to take a cross-validation approach, 
estimating the parameters of the model using a ``training" partition of the data,
then calculating the likelihood of a separate, ``testing" partition of the data given those parameters,
repeating this procedure many times, 
and selecting the model with the significantly best predictive accuracy
(see Section \ref{Xvalidation} for a thorough discussion of our cross-validation procedure).
This approach rewards models with larger values of $K$, which increase the likelihood of the training partition,
up to the point that cluster membership is invoked to model noise specific to the training data, 
at which point predictive accuracy decreases, 
and models with additional clusters are penalized.

>>>>

\paragraph{Choice of cluster number and cross-validation}
There are a number of reasons, both statistical and biological, why
there is no true (or right) number of clusters for real datasets. 
%There are a number of reasons, both statistical and biological, 
%why choosing a single, ``best" number of clusters to describe the data may not be desirable.
%, including ambiguity in what exactly it means for any given value of $K$ to be ``best."
We discuss these reasons further in the Discussion below. In practice
it is useful to explore the implications of clustering at a variety of
different levels (i.e. across multiple $K$) for biological
intepretation and downstream analyses. 
However, we also recognize that in many circumstances it is 
useful to know the number of clusters with highest statistical
support in the (current) dataset.  %, for example to focus su 
%However, we also recognize that in many circumstances 
i%t is either useful or convenient 
%to present only a single set of results from a clustering analysis with a particular number of clusters.
%ideally a number large enough to describe meaningful partitions in the data, 
%but not so large as to be difficult to parse visually.
Toward that end, we have implemented a method for 
statistically comparing \texttt{conStruct} results across different values of $K$.

Several approaches have been used as model choice criteria for the number of
clusters in \gc{population genetic} data, including: 
the use of information theoretic criteria \citep[e.g.][]{ADMIXTURE}, such as AIC or BIC;
comparisons of the likelihood of the data across different values of $K$, 
with various criteria on how to choose a single value (e.g., \citep{Evanno2005});
or comparisons of the marginal likelihood, 
generated either via various approximations \citep[e.g.][]{STRUCTURE}
or via a more rigorous and computationally intensive approach such as thermodynamic integration \citep{verity_nichols2016}
or inference using a Dirichlet process prior \citep{huelsenbeck2007inference}
See \cite{verity_nichols2016} for a discussion of these approaches and comparison
between several methods. %A comprehensive review of these approaches and their various merits is beyond the scope of this paper,
%but

\gc{IS PREDICTION accuracy the right phrase?}
We choose to take a cross-validation approach.  We estimate the
posterior on the parameters of the model using a ``training" partition
of the data, in practice the training set is a random subset of $90\%$
of the loci. We then calculate the likelihood of the
``testing" subset, the remaining $10\%$ of the loci, averaged over the
posterior of the parameters. We repeat this procedure a number of
times, take an average of the relative log-likelihood over the runs at
sumarize each $K$. 
and selecting the model with the significantly best predictive accuracy
(see Section \ref{Xvalidation} for more on our cross-validation procedure).
This approach rewards models with larger values of $K$, which increase
the likelihood of the training partition, as they fit the data better,
but the improvement in the likelihood will plateau (or peak) as above
a certain $K$ we are fitting noise specific to the training data and
so not fitting the testing set any better, 
%up to the point that cluster membership is invoked to model noise specific to the training data, 
%at which point predictive accuracy decreases, 
%and models with additional clusters are penalized.
